Fluoroproj: Title to be decided later

### 2023-11-17 manuscript/manuscript.qmd
started on site map
figures for review are done!
worked on discussion

Possible Journals: Lake and Reservoir Management, PLOS One, PLOS Water, Ecosphere, F1000Research,

@@

Cyanophytes were the dominant taxa by relative biovolume at Curran (0.75), Mashapaug (0.87), and Warwick (0.89); Chrysophytes at Barber (0.39) and Yawgoo (0.44); and Bacillariophytes at Indian (0.41)(Figure 4).  Total biovolume concentrations were highest at Mashapaug (2.12x10^7^), Warwick (1.44x10^7^), and Curran (3.22x10^6^)(Figure 5). Microcystis and Woronchinia were the dominant genera at Mashapaug and Warwick, while Chrysosporum was dominant at Curran (Table 1).

@@

Fig 1. Map of Field Sites

@@

```{r read_table1, echo=FALSE, message=FALSE}
table1 <- read_csv(here::here("data/cleaned_phycotech_cyano_data.csv")) %>%
  select(-family) %>%
  mutate(biovolume = round(biovolume, 1),
         relative_biovolume = round(relative_biovolume, 5))
```

@@

>>>>

<!-- Possible Journals: Lake and Reservoir Management, PLOS One, PLOS Water, Ecosphere, F1000Research, -->

@@

Cyanophytes were the dominant taxa by relative biovolume at Curran (0.75), Mashapaug (0.87), and Warwick (0.89); Chrysophytes at Barber (0.39) and Yawgoo (0.44); and Bacillariophytes at Indian (0.41)(Figure 4).  Total biovolume concentrations were highest at Mashapaug (2.12x10^7^), Warwick (1.44x10^7^), and Curran (3.22x10^6^)(Figure 5). Microcystis was the dominant genus at Mashapaug and Warwick, while Chrysosporum was dominant at Curran (Table 1).

@@

![Fig 1. Map of Field Sites](images/fig1_map.jpg)

@@

```{r read_table1, echo=FALSE, message=FALSE}
library(dplyr)
library(readr)
table1 <- read_csv(here::here("data/cleaned_phycotech_cyano_data.csv")) %>%
  select(-family) %>%
  mutate(biovolume = round(biovolume, 1),
         relative_biovolume = round(relative_biovolume, 5))
```

@@

The results from this study indicate that in-vivo phycocyanin is better related to extracted phycocyanin when compared to the same measurements for chlorophyll.  Similarly, phycocyanin from all instruments was better correlated to cell counts than chlorophyll measurements.  Instruments measuring phycocyanin were better able to handle differences in the phytoplankton community between these study ponds that cause chlorophyll measurements to be less strongly correlated with extracted results.  Both extracted phycocyanin and chlorophyll were strongly correlated with cell counts across the six study ponds suggesting that pigments are indeed a good proxy for cell counts at least for some waterbodies.
The relationship between phycocyanin and all fluorometers was strong.  The Fluorosense had the strongest correlation even with the differences in phytoplankton concentrations and community dynamics between the waterbodies. However, the Fluorosense overestimated the actual concentrations as compared to extracted values (0-35 ug/L vs 0-12.5 ug/L).  Hambdhani et al. also found that the Fluorosense performed well but overestimated chlorophyll a at higher algal concentrations (Hambdhani et al. 2021).  Therefore, using this particular instrument to enforce set criteria across a variety of waterbodies would probably not be recommended without further research. The Algaetorch and Phycoprobe yielded similar measurements and also overestimated phycocyanin concentrations but the output from these instruments is not a direct measure of phycocyanin; instead, these instruments measure the percentage of the chlorophyll that is derived from cyanobacteria.  The Cyanofluor output was in rfus and was not directly comparable to extracted phycocyanin without creating a standard curve for the instrument using known phycocyanin standard concentrations, which was beyond the scope of this study.  A study by Thomson-Laing et al. did convert rfus to ug/L and found strong correlation between cyanobacteria biovolume and phycocyanin concentration for eleven cyanobacteria cultures (Thomson-Laing et al. 2020).  They also found stronger relationships for environmental samples in lakes that were cyanobacteria dominated compared to more diverse lakes.  All of these instruments could be used to effectively monitor for the presence of cyanobacteria by measuring more direct parameters of cyanobacteria than chlorophyll a.
All of the fluorometers exhibited a positive relationship with chlorophyll, but the Algaetorch and Phycoprobe had a much stronger relationship.  Reduced correlation by the Cyanofluor and Trilogy in vivo module was driven by the underestimation of chlorophyll in Mashapaug Pond and possible overestimation in Warwick and Curran.  Phycocyanin concentrations and total biovolume of cyanobacteria were the highest in Mashapaug and it is likely that these high concentrations were underestimating the in-vivo concentrations of chlorophyll (Zamyadi et al . 2012). Therefore, the Algaetorch and Phycoprobe are more successful in properly estimating chlorophyll across a range of different waterbodies, but they are also the most expensive instruments.  A study by Silva et al. found that the Fluoroprobe (a similar instrument to the Phycoprobe) was correlated with cell counts at chlorophyll a concentrations below 100 ug/L but not above, but the present study did not have study ponds that exceed 100 ug/L (Silva et al. 2016).
This study supports the use of extracted chlorophll and phycocyanin as an alternative to cell counts for the study waterbodies.  Currently, in vivo chlorophyll a is typically used for routine monitoring of cyanoHABs, but this study found that phycocyanin to be better correlated with extracted phycocyanin across a variety of instruments. Therefore, measuring in vivo phycocyanin may be a good proxy for cell counts. There are a variety of instruments capable of measuring phycocyanin using in-vivo fluorometry across a spectrum of prices.  We hope this study will encourage further research and consideration of using in-vivo phycocyanin as alternative to cell counts and in situations were extractive techniques are not able to be used.
